---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# _bomrang_: Fetch Australian Government Bureau of Meteorology (BOM) Data

[![Travis-CI Build Status](https://travis-ci.org/ToowoombaTrio/bomrang.svg?branch=master)](https://travis-ci.org/ToowoombaTrio/bomrang)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ToowoombaTrio/bomrang?branch=master&svg=true)](https://ci.appveyor.com/project/ToowoombaTrio/bomrang)
[![Coverage Status](https://img.shields.io/codecov/c/github/ToowoombaTrio/bomrang/master.svg)](https://codecov.io/github/ToowoombaTrio/bomrang?branch=master)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/toowoombatrio/bomrang/commits/master) 
[![minimal R version](https://img.shields.io/badge/R%3E%3D-`r as.character(getRversion())`-brightgreen.svg)](https://cran.r-project.org/)
[![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/) 

Fetches Australian Government Bureau of Meteorology XML précis forecast and
returns a tidy data frame ([Tibble](http://tibble.tidyverse.org)) of the daily 
weather forecast.

Credit for the name, _bomrang_, goes to [Di Cook](http://dicook.github.io), who
suggested it attending the rOpenSci AUUnconf in Brisbane, 2016, when seeing the
[vignette](https://github.com/saundersk1/auunconf16/blob/master/Vignette_BoM.pdf)
that we had assembled during the Unconf.

## Quick start

```{r, eval=FALSE}
if (!require("devtools")) {
  install.packages("devtools", repos = "http://cran.rstudio.com/") 
  library("devtools")
}

devtools::install_github("toowoombatrio/bomrang")
```

## Using _bomrang_

_bomrang_ allows you to fetch forecasts for an individual state or all at once,
the national forecast. To fetch an individual state, simply use the official postal code
for the state for the `state` parameter.
### Using `get_forecast()`

This function only takes one parameter, `state`. The `state` parameter allows the user to select the forecast for just one state or a national forecast. States are specified using the official postal codes, 

**ACT** - Australian Capital Territory  
**NSW** - New South Wales  
**NT** - Northern Territory  
**QLD** - Queensland  
**SA** - South Australia  
**TAS** - Tasmania  
**VIC** - Tasmania  
**WA** - Western Australia  
**AUS** - Australia, returns forecast for all states  

### Results

The function, `get_forecast()` will return a Tibble of the weather forecast for the daily forecast with the following fields,

**aac** - AMOC Area Code, e.g. WA_MW008, a unique identifier for each location  
**start_time_local** - Start of forecast date and time in local TZ  
**end_time_local** - End of forecast date and time in local TZ  
**start_time_utc** - Start of forecast date and time in UTC  
**end_time_utc** - End of forecast date and time in UTC  
**max_temp** - Maximum forecasted temperature (degrees Celsius)  
**min_temp** - Minimum forecasted temperature (degrees Celsius)  
**lower_prcp_limit** - Lower forecasted precipitation limit (millimetres)  
**upper_prcp_limit** - Upper forecasted precipitation limit (millimetres)  
**precis** - Précis forecast (a short summary, less than 30 characters)  
**prob_prcp** - Probability of precipitation (percent)  
**location** - Named location for forecast  
**state** - State name (postal code abbreviation)  
**lon** - Longitude of named location (decimal Degrees)  
**lat** - Latitude of named location (decimal Degrees)  
**elev** - Elevation of named location (metres)  

### Examples

Following is an example fetching the forecast for Queensland.

```{r, eval=TRUE}
library("bomrang")

QLD_forecast <- get_forecast(state = "QLD")
QLD_forecast
```

### Using `get_bulletin()`

This function only takes one parameter, `state`. The `state` parameter allows the user to select the bulletin for just one state or a national forecast. States are specified using the official postal codes, 

**ACT** - Australian Capital Territory  
**NSW** - New South Wales  
**NT** - Northern Territory  
**QLD** - Queensland  
**SA** - South Australia  
**TAS** - Tasmania  
**VIC** - Tasmania  
**WA** - Western Australia  
**AUS** - Australia, returns forecast for all states  

### Results

The function, `get_bulletin()` will return a Tibble of the agriculture bulletin for Queensland with the following fields,

**obs-time-utc** - Observation time (Time in UTC)  
**time-zone** - Time zone for observation  
**site** - Unique BOM identifier for each station  
**name** - BOM station name  
**r** - Rain to 9am (millimetres). \strong{Trace will be reported as 0.01  
**tn** - Minimum temperature (Degrees Celsius)  
**tx** - Maximum temperature (Degrees Celsius)  
**twd** - Wetbulb depression (Degress Celsius)  
**ev** - Evaporation (milimetres)  
**tg** - Terrestrial minimum temperature (Degress Celsius)  
**sn** - Sunshine (Hours)  
**t5** - 5cm temperature (Celsius)  
**t10** - 10cm temperature (Celsius)  
**t20** - 20cm temperature (Celsius)  
**t50** - 50cm temperature (Celsius)  
**t1m** - 1m temperature (Celsius)  
**wr** - Wind run (kilometres)  
**state** - State name (postal code abbreviation)  
**lat** - Latitude (Decimal degrees)  
**lon** - Longitude (Decimal degrees)  

```{r, eval=TRUE}
library("bomrang")

QLD_forecast <- get_bulletin(state = "QLD")
QLD_forecast
```


## Meta

* Please [report any issues or bugs](https://github.com/ToowoombaTrio/bomrang/issues).  
* License: MIT  
* To cite _bomrang_, please use:  
  Sparks A and Pembleton K (2017). _bomrang: Fetch Australian Government Bureau of Meteorology Weather Data_. R package version 0.0.1-1,
<URL: https://github.com/ToowoombaTrio/bomrang>.  
* Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.  

## References

[Australian Bureau of Meteorology (BOM) Weather Data Services](http://www.bom.gov.au/catalogue/data-feeds.shtml)
