---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# _BOMRang_: Fetch Australian Government Bureau of Meteorology (BOM) Data

[![Travis-CI Build Status](https://travis-ci.org/ToowoombaTrio/BOMRang.svg?branch=master)](https://travis-ci.org/ToowoombaTrio/BOMRang)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/ToowoombaTrio/BOMRang?branch=master&svg=true)](https://ci.appveyor.com/project/ToowoombaTrio/BOMRang)
[![Coverage Status](https://img.shields.io/codecov/c/github/ToowoombaTrio/BOMRang/master.svg)](https://codecov.io/github/ToowoombaTrio/BOMRang?branch=master)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/toowoombatrio/BOMRang/commits/master) 
[![minimal R version](https://img.shields.io/badge/R%3E%3D-`r as.character(getRversion())`-brightgreen.svg)](https://cran.r-project.org/)
[![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/) 

Fetches Australian Government Bureau of Meteorology XML précis forecast and
returns a tidy data frame ([Tibble](http://tibble.tidyverse.org)) of the daily 
weather forecast.

Credit for the name, _BOMRang_, goes to [Di Cook](http://dicook.github.io), who
suggested it attending the rOpenSci AUUnconf in Brisbane, 2016, when seeing the
[vignette](https://github.com/saundersk1/auunconf16/blob/master/Vignette_BoM.pdf)
that we had assembled during the Unconf.

## Quick start

```{r, eval=FALSE}
if (!require("devtools")) {
  install.packages("devtools", repos = "http://cran.rstudio.com/") 
  library("devtools")
}

devtools::install_github("toowoombatrio/BOMRang")
```

## Using _BOMRang_

_BOMRang_ allows you to fetch forecasts for an individual state or all at once,
the national forecast. To fetch an individual state, simply use the official postal code
for the state for the `state` parameter.

Fetch the forecast for Queensland
```{r, eval=TRUE}
library("BOMRang")

QLD_forecast <- get_forecast(state = "QLD")
QLD_forecast
```

If you want the national forecast for all areas BOM offers, use the official 
three letter ISO code abbreviation for Australia, AUS, as the `state` parameter.

```{r, eval=FALSE}
BOM_forecast <- get_forecast(state = "AUS")
```

## Results

The function, `get_forecast()` will return a Tibble of the weather forecast for  daily forecast with the following fields.

* **aac** - AMOC Area Code, e.g. WA_MW008, a unique identifier for each location  
* **date** - Date (YYYY-MM-DD)  
* **max_temp** - Maximum forecasted temperature (degrees Celsius)  
* **min_temp** - Minimum forecasted temperature (degrees Celsius)  
* **lower_prcp_limit** - Lower forecasted precipitation limit (millimetres)  
* **upper_prcp_limit** - Upper forecasted precipitation limit (millimetres)  
* **precis** - Précis forecast (a short summary, less than 30 characters)  
* **prob_prcp** - Probability of precipitation (percent)  
* **location** - Named location for forecast  
* **state** - State name (postal code abbreviation)  
* **lon** - Longitude of named location (decimal Degrees)  
* **lat** - Latitude of named location (decimal Degrees)  
* **elev** - Elevation of named location (metres)  


## Meta

* Please [report any issues or bugs](https://github.com/ToowoombaTrio/BOMRang/issues).  
* License: MIT  
* To cite _BOMRang_, please use:  
  Sparks A and Pembleton K (2017). _BOMRang: Fetch Australian Government Bureau of Meteorology Weather Data_. R package version 0.0.1-1,
<URL: https://github.com/ToowoombaTrio/BOMRang>.  
* Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.  

## References

[Australian Bureau of Meteorology (BOM) Weather Data Services](http://www.bom.gov.au/catalogue/data-feeds.shtml)
