---
title: "Get BoM Précis Forecast Town Names"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get BoM Forecast Town Names and Geographic Locations

BoM maintains a shapefile of forecast town names and their geographic locations.
For ease, we'll just use the .dbf file part of the shapefile to extract AAC
codes that can be used to add lat/lon values to the forecast `data.frame` that 
`get_precis_forecast()` returns.

```{r get_forecast_towns}
  utils::download.file(
    "ftp://ftp.bom.gov.au/anon/home/adfd/spatial/IDM00013.dbf",
    destfile = paste0(tempdir(), "AAC_codes.dbf"),
    mode = "wb"
  )

  AAC_codes <-
    foreign::read.dbf(paste0(tempdir(), "AAC_codes.dbf"), as.is = TRUE)
  AAC_codes <- AAC_codes[, c(2:3, 7:9)]
```

Save the stations to disk for use in the R package.

```{r create_data}
 if (!dir.exists("../inst/extdata")) {
      dir.create("../inst/extdata", recursive = TRUE)
    }

  save(AAC_codes, file = "../inst/extdata/AAC_codes.rda",
     compress = "bzip2")
```

## Session Info
```{r session_info}
devtools::session_info()
```
