---
title: "Create BOM Radar Location Database"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get BOM Radar Locations

BOM maintains a shapefile of radar site names and their geographic locations.
For ease, we'll just use the .dbf file part of the shapefile to extract the 
product codes and radar locations. The file is available from BOM's anonymous
FTP server with spatial data <ftp://ftp.bom.gov.au/anon/home/adfd/spatial/>,
specifically the DBF file portion of a shapefile, 
<ftp://ftp.bom.gov.au/anon/home/adfd/spatial/IDR00007.dbf>

```{r get_radarlocs}
curl::curl_download(
  "ftp://ftp.bom.gov.au/anon/home/adfd/spatial/IDR00007.dbf",
  destfile = paste0(tempdir(), "radar_locations.dbf"),
  mode = "wb",
  quiet = TRUE
)

radar_locations <-
  foreign::read.dbf(paste0(tempdir(), "radar_locations.dbf"), as.is = TRUE)

radar_locations$LocationID <-
  ifelse(
    test = nchar(radar_locations$LocationID) == 1,
    yes = paste0("0", radar_locations$LocationID),
    no = radar_locations$LocationID
  )

data.table::setDT(radar_locations)
data.table::setkey(radar_locations, "Name")

str(radar_locations)
```

Save the radar locations to disk for use in the R package.

```{r create_data}
if (!dir.exists("../inst/extdata")) {
  dir.create("../inst/extdata", recursive = TRUE)
}

save(radar_locations,
     file = "../inst/extdata/radar_locations.rda",
     compress = "bzip2"
)
```

## Session Info
```{r session_info}
sessioninfo::session_info()
```
