---
title: "Get BOM Marine Zones"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get BOM Forecast Marine Zones

BOM maintains a shapefile of forecast marine zone names and their geographic
locations. For ease, we'll just use the .dbf file part of the shapefile to
extract AAC codes that can be used to add locations to the forecast `data.frame`
that `get_coastal_forecast()` returns. The file is available from BOM's
anonymous FTP server with spatial data
<ftp://ftp.bom.gov.au/anon/home/adfd/spatial/>, specifically the DBF file
portion of a shapefile,
<ftp://ftp.bom.gov.au/anon/home/adfd/spatial/IDM00003.dbf>

```{r get_forecast_towns}
  curl::curl_download(
    "ftp://ftp.bom.gov.au/anon/home/adfd/spatial/IDM00003.dbf",
    destfile = paste0(tempdir(), "marine_AAC_codes.dbf"),
    mode = "wb",
    quiet = TRUE
  )

  marine_AAC_codes <-
    foreign::read.dbf(paste0(tempdir(), "marine_AAC_codes.dbf"), as.is = TRUE)
  
  marine_AAC_codes <- marine_AAC_codes[, c(1,3,4,5,6,7)]
```

Save the marine zones to disk for use in the R package.

```{r create_data}
 if (!dir.exists("../inst/extdata")) {
      dir.create("../inst/extdata", recursive = TRUE)
    }

  save(marine_AAC_codes, file = "../inst/extdata/marine_AAC_codes.rda",
     compress = "bzip2")
```

## Session Info
```{r session_info}
sessioninfo::session_info()
```
