---
title: "Create BOM Marine Zones Database"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get BOM Forecast Marine Zones

BOM maintains a shapefile of forecast marine zone names and their geographic
locations. For ease, we'll just use the .dbf file part of the shapefile to
extract AAC codes that can be used to add locations to the forecast `data.table`
that `get_coastal_forecast()` returns. The file is available from BOM's
anonymous FTP server with spatial data
<ftp://ftp.bom.gov.au/anon/home/adfd/spatial/>, specifically the DBF file
portion of a shapefile,
<ftp://ftp.bom.gov.au/anon/home/adfd/spatial/IDM00003.dbf>

```{r get_forecast_towns}
curl::curl_download(
  "ftp://ftp.bom.gov.au/anon/home/adfd/spatial/IDM00003.dbf",
  destfile = paste0(tempdir(), "marine_AAC_codes.dbf"),
  mode = "wb",
  quiet = TRUE
)

marine_AAC_codes <-
  foreign::read.dbf(paste0(tempdir(), "marine_AAC_codes.dbf"), as.is = TRUE)

# convert names to lower case for consistency with bomrang output
names(marine_AAC_codes) <- tolower(names(marine_AAC_codes))

# reorder columns
marine_AAC_codes <- marine_AAC_codes[, c(1, 3, 4, 5, 6, 7)]

data.table::setDT(marine_AAC_codes)
data.table::setkey(marine_AAC_codes, "aac")

str(marine_AAC_codes)
```

Save the marine zones to disk for use in bomrang.

```{r create_data}
if (!dir.exists("../inst/extdata")) {
  dir.create("../inst/extdata", recursive = TRUE)
}

save(marine_AAC_codes, file = "../inst/extdata/marine_AAC_codes.rda",
     compress = "bzip2")
```

## Session Info
```{r session_info}
sessioninfo::session_info()
```
