---
title: "Create BOM Précis Forecast Town Names Database"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get BOM Forecast Town Names and Geographic Locations

BOM maintains a shapefile of forecast town names and their geographic locations.
For ease, we'll just use the .dbf file part of the shapefile to extract AAC
codes that can be used to add lat/lon values to the forecast `data.table` that 
`get_precis_forecast()` returns. The file is available from BOM's anonymous
FTP server with spatial data <ftp://ftp.bom.gov.au/anon/home/adfd/spatial/>,
specifically the DBF file portion of a shapefile, 
<ftp://ftp.bom.gov.au/anon/home/adfd/spatial/IDM00013.dbf>

```{r get_forecast_towns}
curl::curl_download(
  "ftp://ftp.bom.gov.au/anon/home/adfd/spatial/IDM00013.dbf",
  destfile = paste0(tempdir(), "AAC_codes.dbf"),
  mode = "wb",
  quiet = TRUE
)

AAC_codes <-
  foreign::read.dbf(paste0(tempdir(), "AAC_codes.dbf"), as.is = TRUE)

# convert names to lower case for consistency with bomrang output
names(AAC_codes) <- tolower(names(AAC_codes))

# reorder columns
AAC_codes <- AAC_codes[, c(2:3, 7:9)]

data.table::setDT(AAC_codes)
data.table::setnames(AAC_codes, c(2, 5), c("town", "elev"))
data.table::setkey(AAC_codes, "aac")
str(AAC_codes)
```

Save the stations to disk for use in the R package.

```{r create_data}
if (!dir.exists("../inst/extdata")) {
  dir.create("../inst/extdata", recursive = TRUE)
}

save(AAC_codes,
     file = "../inst/extdata/AAC_codes.rda",
     compress = "bzip2"
)
```

## Session Info
```{r session_info}
sessioninfo::session_info()
```
