---
title: "Create BOM Radar Location Database"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r color, echo = FALSE, results='asis'}
# crayon needs to be explicitly activated in Rmd
options(crayon.enabled = TRUE)
# Hooks needs to be set to deal with outputs
# thanks to fansi logic
old_hooks <- fansi::set_knit_hooks(knitr::knit_hooks, 
                                   which = c("output", "message", "error"))
```

## Get BOM Radar Locations

BOM maintains a shapefile of radar site names and their geographic locations.
For ease, we'll just use the .dbf file part of the shapefile to extract the product codes and radar locations.
The file is available from BOM's anonymous FTP server with spatial data <ftp://ftp.bom.gov.au/anon/home/adfd/spatial/>, specifically the DBF file portion of a shapefile, 
<ftp://ftp.bom.gov.au/anon/home/adfd/spatial/IDR00007.dbf>.

```{r get_radarlocs}
curl::curl_download(
  "ftp://ftp.bom.gov.au/anon/home/adfd/spatial/IDR00007.dbf",
  destfile = paste0(tempdir(), "radar_locations.dbf"),
  mode = "wb",
  quiet = TRUE
)

new_radar_locations <-
  foreign::read.dbf(paste0(tempdir(), "radar_locations.dbf"), as.is = TRUE)

new_radar_locations$LocationID <-
  ifelse(
    test = nchar(new_radar_locations$LocationID) == 1,
    yes = paste0("0", new_radar_locations$LocationID),
    no = new_radar_locations$LocationID
  )

data.table::setDT(new_radar_locations)
data.table::setkey(new_radar_locations, "Name")

str(new_radar_locations)
```

## Show Changes from Last Release

To ensure that the data being compared is from the most recent release, reinstall _bomrang_ from CRAN.

```{r diff-codes}
install.packages("bomrang", repos = "http://cran.us.r-project.org")

load(system.file("extdata", "radar_locations.rda", package = "bomrang"))

(
  radar_location_changes <-
    diffobj::diffPrint(new_radar_locations, radar_locations)
)
```

# Save Radar Stations Data and Changes

Save the radar stations' metadata and changes to disk for use in _bomrang_.

```{r create_data}
if (!dir.exists("../inst/extdata")) {
  dir.create("../inst/extdata", recursive = TRUE)
}

radar_locations <- new_radar_locations

save(radar_locations,
     file = "../inst/extdata/radar_locations.rda",
     compress = "bzip2")

save(radar_location_changes,
     file = "../inst/extdata/radar_location_changes.rda",
     compress = "bzip2")
```

## Session Info
```{r session_info}
sessioninfo::session_info()
```
