---
title: "_bomrang_: Fetch Australian Government Bureau of Meteorology (BOM) Data"
author: "Adam H Sparks"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bomrang}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

_bomrang_ provides functions for interacting with [Australian Bureau of Meteorology (BOM) Weather Data Services](http://www.bom.gov.au/catalogue/data-feeds.shtml) forecasts. BOM serves several types of data data as XML and JSON files. This package fetches these files, parses them and return a tidy dataframe.

## Using _bomrang_

Several functions are provided by _bomrang_. The three main functions are `get_precis_forecast()`, which creates a data frame of the daily précis forecast from BOM, `get_ag_bulletin()`, which creates a data frame of agriculture bulletin information, and `get_current_weather()` which returns the last 72 hours of observations at a given location. 
Other functions include `update_locations()`, which updates an internal database of forecast locations distributed with the package, and
`sweep_for_stations()` which returns the nearest weather stations to a point in Australia.

## Using `get_precis_forecast()`

This function only takes one parameter, `state`. The `state` parameter allows the user to select the forecast for just one state or a national forecast. States are specified using the official postal codes, 

* **ACT** - Australian Capital Territory  
* **NSW** - New South Wales  
* **NT** - Northern Territory  
* **QLD** - Queensland  
* **SA** - South Australia  
* **TAS** - Tasmania  
* **VIC** - Victoria  
* **WA** - Western Australia  
* **AUS** - Australia, returns national forecast including all states  

### Results

The function, `get_precis_forecast()` will return a [tibble](https://CRAN.R-project.org/package=tibble) of the weather forecast for the daily forecast with the following fields,

* **aac** - AMOC Area Code, _e.g._ WA_MW008, a unique identifier for each location  
* **start_time_local** - Start of forecast date and time in local TZ  
* **end_time_local** - End of forecast date and time in local TZ  
* **start_time_utc** - Start of forecast date and time in UTC  
* **end_time_utc** - End of forecast date and time in UTC  
* **max_temp** - Maximum forecasted temperature (degrees Celsius)  
* **min_temp** - Minimum forecasted temperature (degrees Celsius)  
* **lower_prcp_limit** - Lower forecasted precipitation limit (millimetres)  
* **upper_prcp_limit** - Upper forecasted precipitation limit (millimetres)  
* **precis** - Précis forecast (a short summary, less than 30 characters)  
* **prob_prcp** - Probability of precipitation (percent)  
* **location** - Named location for forecast  
* **state** - State name (postal code abbreviation)  
* **lon** - Longitude of named location (decimal degrees)  
* **lat** - Latitude of named location (decimal degrees)  
* **elev** - Elevation of named location (metres)  

### Examples

Following is an example fetching the forecast for Queensland.

```{r, eval=TRUE}
library("bomrang")

QLD_forecast <- get_precis_forecast(state = "QLD")
QLD_forecast
```

## Using `get_ag_bulletin()`

This function only takes one parameter, `state`. The `state` parameter allows the user to select the bulletin for just one state or a national forecast. States are specified using the official postal codes, 

* **ACT** - Australian Capital Territory  
* **NSW** - New South Wales  
* **NT** - Northern Territory  
* **QLD** - Queensland  
* **SA** - South Australia  
* **TAS** - Tasmania  
* **VIC** - Victoria  
* **WA** - Western Australia  
* **AUS** - Australia, returns bulletin for all states  

### Results

The function, `get_ag_bulletin()` will return a [tibble](https://CRAN.R-project.org/package=tibble) of the agriculture bulletin with the following fields,

* **obs-time-utc** - Observation time (Time in UTC)  
* **time-zone** - Time zone for observation  
* **site** - Unique BOM identifier for each station  
* **name** - BOM station name  
* **r** - Rain to 9am (millimetres). _Trace will be reported as 0.01_  
* **tn** - Minimum temperature (degrees Celsius)  
* **tx** - Maximum temperature (degrees Celsius)  
* **twd** - Wetbulb depression (degress Celsius)  
* **ev** - Evaporation (millimetres)  
* **tg** - Terrestrial minimum temperature (degress Celsius)  
* **sn** - Sunshine (Hours)  
* **t5** - 5cm soil temperature (Celsius)  
* **t10** - 10cm soil temperature (Celsius)  
* **t20** - 20cm soil temperature (Celsius)  
* **t50** - 50cm soil temperature (Celsius)  
* **t1m** - 1m soil temperature (Celsius)  
* **wr** - Wind run (kilometres)  
* **state** - State name (postal code abbreviation)  
* **lat** - Latitude (decimal degrees)  
* **lon** - Longitude (decimal degrees)  

```{r, eval=TRUE}
library("bomrang")

QLD_bulletin <- get_ag_bulletin(state = "QLD")
QLD_bulletin
```

## Using `get_current_weather()` 
The function takes one of two arguments: `station_name` and `latlon`, returning the current weather observations (and the observations of the last 72 hours) for the given location.

If `station_name` is used, the weather observations for the last 72 hours are returned for that station.
If the string provided is ambiguous, the function returns an observation for one of the possible stations and emits a warning to offer unambiguous station names. 

If `latlon` is used, the observations returned are from the station nearest to that latitude-longitude coordinate. 
The function also emits a message, to tell the user which station was used.

The table returned will have different fields depending on the station that is selected. 



## Using `update_locations()`

This function downloads the latest précis forecast locations from the BOM server and updates _bomrang_'s internal database of forecast locations. There is no need to use this unless you know that a forecast location exists in a more current version of the BOM forecast location database that is not available in the database distributed with _bomrang_. This database is distributed with the package to make the process faster when fetching the forecast.

```{r, eval=FALSE}
update_locations()
```

## References

[Australian Bureau of Meteorology (BOM) Weather Data Services](http://www.bom.gov.au/catalogue/data-feeds.shtml)

[Australian Bureau of Meteorology (BOM) Weather Data Services Agriculture Bulletins](http://www.bom.gov.au/catalogue/observations/about-agricultural.shtml)

[Australian Bureau of Meteorology (BOM) Weather Data Services Observation of Rainfall](http://www.bom.gov.au/climate/how/observations/rain-measure.shtml)