---
title: "bomrang"
author: "Adam H Sparks"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bomrang}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

_bomrang_ provides functions for interacting with [Australian Bureau of Meteorology (BOM) Weather Data Services](http://www.bom.gov.au/catalogue/data-feeds.shtml) forecasts. BOM serves several types of data data as XML and JSON files. This package fetches these files, parses them and return a tidy data frame.

## Using _bomrang_

Several functions are provided by _bomrang_.The main functionality of _bomrang_ is provided through three functions, `get_precis_forecast()`, which retrieves the précis (short) forecast; `get_current_weather()`, which fetches the current weather from a given station; and `get_ag_bulletin()`, which retrieves the agriculture bulletin. Other functions include `update_forecast_locations()`, which updates an internal database of forecast locations distributed with the package, and `sweep_for_stations()` which returns the nearest weather stations to a point in Australia.

## Using `get_precis_forecast`

This function only takes one parameter, `state`. The `state` parameter allows the user to select the forecast for just one state or a national forecast. States or territories are specified using the official postal codes:

* **ACT** - Australian Capital Territory  
* **NSW** - New South Wales  
* **NT** - Northern Territory  
* **QLD** - Queensland  
* **SA** - South Australia  
* **TAS** - Tasmania  
* **VIC** - Victoria  
* **WA** - Western Australia  
* **AUS** - Australia, returns national forecast including all states.

### Results

The function, `get_precis_forecast()`, will return a data frame of the weather forecast for the daily forecast with the following fields:

* **aac** - AMOC Area Code, _e.g._ WA_MW008, a unique identifier for each location  
* **start_time_local** - Start of forecast date and time in local TZ  
* **end_time_local** - End of forecast date and time in local TZ  
* **UTC_offset** -  Hours offset from difference in hours and minutes from Coordinated Universal Time (UTC) for `start_time_local` and `end_time_local`
* **start_time_utc** - Start of forecast date and time in UTC  
* **end_time_utc** - End of forecast date and time in UTC  
* **max_temp** - Maximum forecasted temperature (degrees Celsius)  
* **min_temp** - Minimum forecasted temperature (degrees Celsius)  
* **lower_prcp_limit** - Lower forecasted precipitation limit (millimetres)  
* **upper_prcp_limit** - Upper forecasted precipitation limit (millimetres)  
* **precis** - Précis forecast (a short summary, less than 30 characters)  
* **prob_prcp** - Probability of precipitation (percent)  
* **location** - Named location for forecast  
* **state** - State name (postal code abbreviation)  
* **lon** - Longitude of named location (decimal degrees)  
* **lat** - Latitude of named location (decimal degrees)  
* **elev** - Elevation of named location (metres).

### Example

Following is an example fetching the forecast for Queensland.

```{r, eval=TRUE}
library("bomrang")

QLD_forecast <- get_precis_forecast(state = "QLD")
head(QLD_forecast)
```

## Using `get_ag_bulletin`

`get_ag_bulletin()` only takes one parameter, `state`. The `state` parameter allows the user to select the bulletin for just one state or a national forecast. States or territories are specified using the official postal codes.

* **NSW** - New South Wales  
* **NT** - Northern Territory  
* **QLD** - Queensland  
* **SA** - South Australia  
* **TAS** - Tasmania  
* **VIC** - Victoria  
* **WA** - Western Australia  
* **AUS** - Australia, returns bulletin for all states/territories.

### Results

The function, `get_ag_bulletin()`, will return a data frame of the agriculture bulletin with the following fields:

* **obs-time-local** - Observation time  
* **obs-time-utc** - Observation time (time in UTC)  
* **time-zone** - Time zone for observation  
* **site** - Unique BOM identifier for each station  
* **dist** - BOM rainfall district  
* **station** - BOM station name  
* **start** - Year data collection starts  
* **end** - Year data collection ends (will always be current)  
* **state** - State name (postal code abbreviation)  
* **lat** - Latitude (decimal degrees)  
* **lon** - Longitude (decimal degrees)  
* **elev** - Station elevation (metres)  
* **bar_ht** - Bar height (metres)  
* **WMO** - World Meteorlogical Organization number (unique ID used worldwide)
* **r** - Rain to 9am (millimetres). _Trace will be reported as 0.01_  
* **tn** - Minimum temperature (degrees Celsius)  
* **tx** - Maximum temperature (degrees Celsius)  
* **twd** - Wet bulb depression (degrees Celsius)  
* **ev** - Evaporation (millimetres)  
* **tg** - Terrestrial minimum temperature (degrees Celsius)  
* **sn** - Sunshine (hours)  
* **t5** - 5cm soil temperature (degrees Celsius)  
* **t10** - 10cm soil temperature (degrees Celsius)  
* **t20** - 20cm soil temperature (degrees Celsius)  
* **t50** - 50cm soil temperature (degrees Celsius)  
* **t1m** - 1m soil temperature (degrees Celsius)  
* **wr** - Wind run (kilometres)  
* **state** - State name (postal code abbreviation)  
* **lat** - Latitude (decimal degrees)  
* **lon** - Longitude (decimal degrees).

### Example

Following is an example fetching the ag bulletin for Queensland.

```{r, eval=TRUE}
library("bomrang")

QLD_bulletin <- get_ag_bulletin(state = "QLD")
head(QLD_bulletin)
```

## Using `get_current_weather` 
`get_current_weather()` takes one of two arguments: `station_name` and `latlon`, returning the current weather observations (and the observations of the last 72 hours) for the given location.

If `station_name` is used, the weather observations for the last 72 hours are returned for that station.
If the string provided is ambiguous, the function returns an observation for one of the possible stations and emits a warning to offer unambiguous station names. 

If `latlon` is used, the observations returned are from the station nearest to that latitude-longitude coordinate. The function also emits a message, to tell the user which station was used.

### Results
The table returned will have different fields depending on the station that is selected. 

### Example
Following is an example fetching the current weather for Melbourne.

```{r, eval=TRUE}
library("bomrang")

Melbourne_weather <- get_current_weather("Melbourne (Olympic Park)")
head(Melbourne_weather)
```

## Other functions in _bomrang_

### Using `sweep_for_stations`
`sweep_for_stations()` only takes one parameter, `latlon`, a length-2 numeric vector. By default, Canberra (approximately).

#### Results
This function will search for weather stations and return a data frame of all weather stations (in this package) sorted by distance from `latlon`, ascending. The fields in the data frame are:

**name** - station name  
**Lat** - latitude (decimal degrees)  
**Lon** - longitude (decimal degrees)  
**distance** - distance from provided `latlon` value (kilometres).

#### Example  

Following is an example sweeping for stations starting with Canberra.  

```{r, eval = TRUE}
# Show only the first ten stations in the list
head(sweep_for_stations(latlon = c(-35.3, 149.2)), 10)
```


### Using `update_forecast_locations`

`update_forecast_locations()` downloads the latest précis forecast locations from the BOM server and updates _bomrang_'s internal database of forecast locations. There is no need to use this unless you know that a forecast location exists in a more current version of the BOM forecast location database that is not available in the database distributed with _bomrang_. This database is distributed with the package to make the process faster when fetching the forecast.

#### Example

Following is an example updating the précis forecast locations internal database.

```{r, eval=FALSE}
update_forecast_locations()
```

### Using `update_station_locations`

`update_station_locations()` downloads the latest station locations and metadata and updates bomrang's internal databases that support the use of `get_current_weather()` and `get_ag_bulletin()`. There is no need to use this unless you know that a station exists in BOM's database that is not available in the databases distributed with _bomrang_

#### Example

Following is an example updating the précis forecast locations internal database.

```{r, eval=FALSE}
update_station_locations()
```

## References

[Australian Bureau of Meteorology (BOM) Weather Data Services](http://www.bom.gov.au/catalogue/data-feeds.shtml)

[Australian Bureau of Meteorology (BOM) Weather Data Services Agriculture Bulletins](http://www.bom.gov.au/catalogue/observations/about-agricultural.shtml)

[Australian Bureau of Meteorology (BOM) Weather Data Services Observation of Rainfall](http://www.bom.gov.au/climate/how/observations/rain-measure.shtml)